---
title: "Dixmoor Municipality Projects"
params: 
  municipality: "Dixmoor"
format: 
  html:
    toc: true
    toc-location: left
    fig-cap-location: margin
    tbl-cap-location: margin
    code-fold: true
    link-external-newwindow: true
    df-print: paged
    message: false

---

[Township directory](../townships.qmd) · [Municipality directory](../municipalities.qmd) · [Search PINs](../pins.qmd)

```{r, message=FALSE, warning=FALSE}
#| label: setup
#| echo: false

library(tidyverse)
library(sf)
library(stringr)
library(readxl)
library(ggplot2)
library(ptaxsim)

# ------------------------------------------------------------------
# Data inputs
# ------------------------------------------------------------------

comm_ind <- read_csv("../Merriman RA/ptax/Output/comm_ind_PINs_2011to2022_timeseries.csv") %>% 
  filter(year == 2022) |> 
  select(-year) |>
    mutate(pin10 = str_sub(pin, 1, 10))


geo_names <- comm_ind |> select(pin, clean_name, Triad, Township)

ptaxsim_db_conn <- DBI::dbConnect(RSQLite::SQLite(), "../Merriman RA/ptax/ptaxsim.db/ptaxsim-2023.0.0.db")


# manually coded projects
keypins <- read_xlsx("../Merriman RA/ptax/Output/projects_checked_MAINFILE.xlsx") |>
  mutate(keypin = main_keypin)

valuation_wksts <- read_csv("../Merriman RA/ptax/Output/keypins_from_methodwkshts.csv")

```


```{r}
#| label: fn-analyze-municipality
#| echo: false

# ------------------------------------------------------------------
# Core function: run analysis for a single municipality
# ------------------------------------------------------------------
analyze_municipality <- function(municipality_name,
                             year = 2022,
                             comm_ind_df = comm_ind,
                             keypins_df = keypins) {
  
  # IMPORTANT:
  # Change `Township` below to whatever column in comm_ind actually holds the township name,
  # e.g. `twnshp`, `township`, or `clean_name`.
  muni_pins <- comm_ind_df |>
    dplyr::filter(clean_name == municipality_name) |>
    dplyr::mutate(pin10 = stringr::str_sub(pin, 1, 10))
  
  
  pins_geo <- lookup_pin10_geometry(year = year, 
                                    pin10 = muni_pins$pin10) %>%
    st_as_sf(wkt = "geometry", crs = 4326)
  
  points <- st_as_sf(pins_geo, coords = c("longitude", "latitude"), crs = 4326)
  
  parcels <- st_as_sf(pins_geo, wkt = "geometry", crs = 4326)
  
  invalid_geom <- st_is_valid(parcels)
#   table(invalid_geom)
  
  parcels$geometry <- st_make_valid(parcels$geometry)
  
  # Filter out invalid geometries
  parcels <- parcels[st_is_valid(parcels), ]
  
  
  ggplot(parcels) +
    geom_sf() +
    theme_minimal() + theme(legend.position = "none")
  
  
  # 2. Join parcels to keypins
  parcel_comparison <- muni_pins |>
    left_join(parcels, by = "pin10") |>
    dplyr::left_join(keypins_df) |>
    dplyr::select(pin10, keypin, dplyr::everything())
  
  # 3. Summary by keypin (from parcels)
  # parcel_keypin_summary <- parcel_comparison |>
  #   sf::st_drop_geometry() |>
  #   dplyr::group_by(keypin) |>
  #   dplyr::summarise(
  #     n           = dplyr::n(),
  #     incent_pins = sum(incent_prop == "Incentive", na.rm = TRUE),
  #     .groups     = "drop"
  #   ) |>
  #   dplyr::arrange(dplyr::desc(n))

  # 4. Map by appealid–keypin combo
  parcel_plot <- parcel_comparison |>
    ggplot2::ggplot() +
    ggplot2::geom_sf(ggplot2::aes( geometry = geometry, fill = keypin)) +
    ggplot2::theme_minimal() +
    ggplot2::scale_fill_discrete(na.value = "white") +
    ggplot2::theme(legend.position = "none")

  # 5. Join to comm_ind-level data for this municipality
  pin_info <- comm_ind_df |>
    dplyr::inner_join(
      parcel_comparison,
      # by = "pin10"
    )

  # 6. Class-level summary (pins + distinct keypins)
  class_summary <- pin_info |>
        sf::st_drop_geometry() |>

  #  dplyr::filter(!is.na(keypin) & !is.na(class)) |>
    dplyr::group_by(class) |>
    dplyr::summarise(
      pins    = dplyr::n(),
      keypins = dplyr::n_distinct(keypin),
     # .groups = "drop"
    )

  # 7. Keypin-level summary from comm_ind
  keypin_info_summary <- pin_info |>
    #dplyr::filter(!is.na(keypin)) |>
    dplyr::group_by(keypin) |>
    dplyr::summarise(
      n             = dplyr::n(),
      appellant      = first(appellant),
      fmv            = sum(fmv,    na.rm = TRUE),
      incentive_pins = sum(incent, na.rm = TRUE),
      exempt         = sum(exempt, na.rm = TRUE),
    #  .groups        = "drop"
    ) |>
    dplyr::arrange(dplyr::desc(incentive_pins), dplyr::desc(n))

  list(
    municipality          = municipality_name,
    parcels               = parcels,
    parcel_comparison     = parcel_comparison,
  #  parcel_keypin_summary = parcel_keypin_summary,
    parcel_plot           = parcel_plot,
    pin_info              = pin_info,
    class_summary         = class_summary,
    keypin_info_summary   = keypin_info_summary
  )
}
```
 
```{r}
#| label: run-analysis

#this_municipality <- params$municipality

res <- analyze_municipality(
  municipality_name = params$municipality,
  year          = 2022,     # adjust if you want a different year
  comm_ind_df   = comm_ind,
  keypins_df    = keypins
)

```

## Municipality overview

There are `r paste(comm_ind |> filter(clean_name == params$municipality) |> distinct(pin) |> count())` PINs in this municipality that were commercial or industrial at least once.


```{r}
#| label: municipality-name
#| echo: false
#| layout-ncol: 2

cat("Municipality:", res$municipality)

townships_in_municipality <- comm_ind |>
  filter(clean_name == params$municipality, !is.na(Township), Township != "") |>
  distinct(Township) |>
  arrange(Township)

townships_in_municipality
```

## Parcel map (outline only)

```{r}
#| label: fig-parcels-outline
#| fig.cap: "All commercial & industrial parcels in the municipality."
#| warning: false
#| message: false

ggplot(res$parcels) +
  geom_sf() +
  theme_void() +
  theme(legend.position = "none")
```

## Manually coded keypins – map

```{r}
#| label: fig-keypin-map
#| fig.cap: "Parcels grouped by manually coded keypins (based off of appeal IDs, sale documents and proximity)."
#| warning: false
#| message: false

res$parcel_plot
```


## Class summary (pins vs keypins)

```{r}
#| label: tbl-class-summary
#| tbl-cap: "PINs and projects per property class"
res$class_summary
```

## Keypin-level summary (fmv, incentives, exemptions)

```{r}
#| label: tbl-keypin-info-summary
#| tbl-cap: "Projects in municipality. You can sort each column!"

res$keypin_info_summary |> DT::datatable(rownames = FALSE)
```


There are `r paste(res$parcel_comparison %>% group_by(keypin) %>% summarize(pins_ingroup = n()) |> filter(pins_ingroup > 1) |> n_distinct())` projects with more than 1 pin in them and  `r paste(res$parcel_comparison %>% group_by(keypin) %>% summarize(pins_ingroup = n()) |> n_distinct())` projects overall based on manually grouped keypins from combined data sources.





```{r}
#| tbl-cap: "This many pins were in the valuation worksheets."


valuation_wksts |> filter(pin %in% res$pin_info$pin)
```

There are `r paste(valuation_wksts |> filter(pin %in% res$pin_info$pin) |> count())` PINs in the valuation worksheets with `r paste(valuation_wksts |> filter(pin %in% res$pin_info$pin)|> distinct(keypin) |> count())` distinct keypins.


There were   `r paste(res$parcel_comparison |> distinct(keypin) |> count() ) ` projects overall based on manually grouped keypins from combined data sources.

There were  `r paste(res$parcel_comparison |> distinct(pin) |> count() ) ` PINs within this municipality that had been commercial or industrial at some point in time. 
